---
output:
  rmarkdown::github_document
bibliography: "README-refs.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`biotmle`

[![Travis-CI Build Status](https://travis-ci.org/nhejazi/biotmle.svg?branch=master)](https://travis-ci.org/nhejazi/biotmle)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/nhejazi/biotmle?branch=master&svg=true)](https://ci.appveyor.com/project/nhejazi/biotmle/)
[![Coverage Status](https://img.shields.io/codecov/c/github/nhejazi/biotmle/master.svg)](https://codecov.io/github/nhejazi/biotmle?branch=master)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![BioC status](http://www.bioconductor.org/shields/build/release/bioc/biotmle.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/biotmle)
[![Bioc Time](http://bioconductor.org/shields/years-in-bioc/biotmle.svg)](https://bioconductor.org/packages/release/bioc/html/biotmle.html)
[![Bioc Downloads](http://bioconductor.org/shields/downloads/biotmle.svg)](https://bioconductor.org/packages/release/bioc/html/biotmle.html)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)
[![DOI](https://zenodo.org/badge/65854775.svg)](https://zenodo.org/badge/latestdoi/65854775)
[![JOSS Status](http://joss.theoj.org/papers/02be843d9bab1b598187bfbb08ce3949/status.svg)](http://joss.theoj.org/papers/02be843d9bab1b598187bfbb08ce3949)

> Targeted Learning with Moderated Statistics for Biomarker Discovery

__Author:__ [Nima Hejazi](https://nimahejazi.org)

---

## What's `biotmle`?

`biotmle` is an R package that facilitates biomarker discovery by generalizing
the moderated t-statistic [@smyth2004linear] for use with target parameters that
have asymptotically linear representations [@vdl2011targeted]. The set of
methods implemented rely on the use of targeted maximum likelihood (TML)
estimation to transform biological sequencing data (e.g., microarray, RNA-seq)
based on the (efficient) influence function (EIF) representation of a particular
causal target parameter (e.g., average treatment effect). The transformed data
(rotated into influence function space) are then be subjected to a moderated
test for differences between the statistical estimate of the target parameter
and a hypothesized value of said parameter (usually a null value defined in
relation to the parameter itself). Such an approach provides a valid and
conservative statistical hypothesis test of a statistically estimable target
parameter while controlling the standard error (derived from the variance of the
EIF) such that the error rate of the test is more strongly controlled relative
to testing procedures that do fail moderate the variance estimate
[@hejazi2018variance] while simultaneously avoiding many of the pitfalls that
plague analogous procedures predicated upon correctly specifying linear models
[@smyth2004linear]. Utilities are also provided for performing clustering based
on _supervised distance matrices_, using the EIF-based estimates to draw out
underlying contributions of biomarkers to the target parameter of interest
[@pollard2008supervised].

---

## Installation

For standard use, install from
[Bioconductor](https://bioconductor.org/packages/biotmle) using
[`BiocManager`](https://CRAN.R-project.org/package=BiocManager):

```{r bioc-installation, eval = FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE)) {
  install.packages("BiocManager")
}
BiocManager::install("biotmle")
```

To contribute, install the bleeding-edge _development version_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-master-installation, eval = FALSE}
devtools::install_github("nhejazi/biotmle")
```

Current and prior [Bioconductor](https://bioconductor.org) releases are
available under branches with numbers prefixed by "RELEASE_". For example, to
install the version of this package available via Bioconductor 3.6, use

```{r gh-develop-installation, eval = FALSE}
devtools::install_github("nhejazi/biotmle", ref = "RELEASE_3_6")
```

---

## Example

For details on how to best use the `biotmle` R package, please consult the most
recent [package
vignette](https://bioconductor.org/packages/release/bioc/vignettes/biotmle/inst/doc/exposureBiomarkers.html)
available through the [Bioconductor
project](https://bioconductor.org/packages/biotmle).

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/nhejazi/biotmle/issues).

---

## Contributions

Contributions are very welcome. Interested contributors should consult our
[contribution
guidelines](https://github.com/nhejazi/biotmle/blob/master/CONTRIBUTING.md)
prior to submitting a pull request.

---

## Citation

After using the `biotmle` R package, please cite both of the following:

        @article{hejazi2017biotmle,
          author = {Hejazi, Nima S and Cai, Weixin and Hubbard, Alan E},
          title = {biotmle: Targeted Learning for Biomarker Discovery},
          journal = {The Journal of Open Source Software},
          volume = {2},
          number = {15},
          month = {July},
          year  = {2017},
          publisher = {The Open Journal},
          doi = {10.21105/joss.00295},
          url = {https://doi.org/10.21105/joss.00295}
        }

        @article{hejazi2018+variance,
          url = {https://arxiv.org/abs/1710.05451},
          year = {2018+},
          author = {Hejazi, Nima S and {Kherad-Pajouh}, Sara and {van der
            Laan}, Mark J and Hubbard, Alan E},
          title = {Variance moderation of locally efficient estimators in
            high-dimensional biology}
        }

---

## Related

* [R/`biotmleData`](https://github.com/nhejazi/biotmleData) - R package with
    example experimental data for use with this analysis package.

---

## Funding

The development of this software was supported in part through grants from the
National Institutes of Health: [P42 ES004705-29](https://projectreporter.nih.gov/project_info_details.cfm?aid=9260357&map=y) and [R01 ES021369-05](https://projectreporter.nih.gov/project_info_description.cfm?aid=9210551&icde=37849782&ddparam=&ddvalue=&ddsub=&cr=1&csb=default&cs=ASC&pball=).

---

## License

&copy; 2016-2018 [Nima S. Hejazi](https://nimahejazi.org)

The contents of this repository are distributed under the MIT license. See file
`LICENSE` for details.

---

## References

